---
title: "Week 1"
description: |
  Getting all the software up and running
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

This week, we will familiarize ourselves with a suite of software that will make life as a social scientist so much nicer -- tools for managing your data, your references, and your knowledge. Make sure you've completed everything from the [Getting Started](index.html#getting-started) section on the home page.

## Introductions

On our first day, we will be collecting and analyzing some data about the class itself. Please complete the following survey.[^survey]

[^survey]: If the embedded survey isn't working for some reason, click [here](https://docs.google.com/forms/d/e/1FAIpQLScvfmmeiPeg4Qudq4E7P-Qiyb7h8mF0CrfusOns4hTGoTdOyQ/viewform?usp=sf_link).

```{r survey}
knitr::include_url('https://docs.google.com/forms/d/e/1FAIpQLScvfmmeiPeg4Qudq4E7P-Qiyb7h8mF0CrfusOns4hTGoTdOyQ/viewform?usp=sf_link')
```

The results will be available [here](https://docs.google.com/spreadsheets/d/1BTVFoZQjRNB63Pyg_z-hjGICpP_vlcokRy0-SSGX-10/).

## Reading Assignments

Before Week 2, please read and annotate the following:

- [Healy Preface](https://socviz.co/index.html#preface) (and complete the package installation instructions at the bottom)
- [Teacup Giraffes](https://tinystats.github.io/teacups-giraffes-and-statistics/index.html) (Intro to R module)
- [Healy Chapter 2](https://socviz.co/gettingstarted.html)
- [Healy Chapter 3](https://socviz.co/makeplot.html#makeplot)


Optional:

- [Healy Chapter 1](https://socviz.co/lookatdata.html): A nice review of the evidence on what makes an effective visualization.

## Problem Set

Our first problem set will ensure that all the software is functioning on your machine. If you can complete all of the following steps, then you're in good shape!

1. Open `intro-to-political-methodology.Rproj` and pull the latest updates from the GitHub repository.

2. Open `R/week-01/problem-set-01.Rmd`.

3. Edit the metadata so that you are the author.

4. Add an in-text citation to something from your Zotero library.

5. Make sure you have the `tinytex` package installed. This is the library you need to convert RMarkdown documents into PDFs. To install, enter these two lines into your RStudio console.

```{r install-tinytex, eval = FALSE, echo = TRUE}
install.packages('tinytex')
tinytex::install_tinytex()
```

6. Knit to PDF.

7. Submit the knitted PDF document to [eLC](https://uga.view.usg.edu/d2l/le/content/2356038/Home) by our next class.[^1]

[^1]: All problem sets due by the *following* week's class (so this one is due before our second class). Feel free to work together on problem sets, but I encourage you to type out the solutions yourself instead of copying-pasting from others. You can't learn this stuff without doing it. 